\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}


    \begin{document}

\title{Shape Detection in an Image using Parallelized Traditional Image Analysis Techniques\\
% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used}
% \thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{
    \IEEEauthorblockN{Alan Manuel Loreto Cornídez}
    \IEEEauthorblockA{\textit{College of Electrical and Computer Engineering} \\
    \textit{The University of Arizona}\\
    Tucson, Arizona \\
    aloretocornidez@arizona.edu}
\and
    \IEEEauthorblockN{2\textsuperscript{nd} Rubén Diego Fuentes Guitérrez}
    \IEEEauthorblockA{\textit{College of Electrical and Computer Engineering} \\
    \textit{The University of Arizona}\\
    Tucson, Arizona \\
    rfuentesgtz@arizona.edu}
}

\maketitle


% \begin{abstract}
% % This document is a model and instructions for \LaTeX.
% % This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL\: Do Not Use Symbols, Special Characters, Footnotes, 
% % or Math in Paper Title or Abstract.
% \end{abstract}

\begin{IEEEkeywords}
image analysis, parallelization, shape detection, telemetry
\end{IEEEkeywords}

% \section{Introduction}
% This document is a model and instructions for \LaTeX.
% Please observe the conference page limits. 
\section*{Abstract}
Modern day computer vision applications are frequently implemented using machine learning approaches.
While these implementations can perform very well, the performance is heavily dependent on sufficient and accurate training data.
Due to a lack of adequate training data, the Arizona Autonomous Vehicles Club (AZA) decided to implement the generalized hough transform to detect shapes in a live video feed from an unmanned aerial system (UAS).
The Hough transform is computationally intensive and since real-time performance is required, a serial approach may not have the execution speed necessary for the application.
Image processing techniques include matrix multiplication and convolution operations which are highly parallelizable.
Therefore, the algorithm was parallelized and implemented on a graphics processing unit (GPU).
Performance profiling was done on both machine learning and traditional approaches where execution time and accuracy were compared.


% If training datasets are not available, not large enough, or not an accurate representation of the application, performance of the implementation model can suffer.
%The generalized hough transform\cite{BALLARD1981111} is used to detect the shapes present in an image.



\section*{Introduction}

The Arizona Autonomous Vehicles Club (AZA) must use computer vision techniques to detect shapes and emergent objects in a video stream.
The current implementation uses a machine learning approach to accomplish this.
There are many benefits to using a trained machine learning model, such as the speed of execution after the model has been trained.
However, there are a few issues that arise to warrant the possibility of a traditional image analysis approach.

First and formost, machine learning implementations require a substantial amount of data in order to train the algorithm and result in satisfactory performance. 
The data that is being used to train the machine learning algorithm must also be similar to the application that will make use of the algorithm. 
The AZA club, unfortunately, does not have neither application specific training data or substantial data due constantly evolving system parameters. 

For these reasons, it has come into the consideration of the club to implement a shape detection algorithm utilizing a traditional image analysis approach.
To accomplish this task, the AZA would like to implement a shape detection algorithm capable of processing a video stream of at least 15 frames per second and detect up to 13 polygonal shapes on a modern consumer grade GeForce RTX series card.
% Profiling of the shape detection algorithm will allow us to determine how much computing power is necessary to apply traditional image analysis techniques to our application.
The generalized hough transform\cite{BALLARD1981111} is used to detect the shapes present in an image.
In recent history, most image analysis applications for computer vision have been implemented using machine learning, specifically, neural networks. 
This project aims to implement traditional digital image analysis techniques to detect shapes in an image.
The generalized hough transform shall be implemented in a parallelized manner by utilizing a graphics processing unit (GPU).
be properly informed when selecting how to proceed with their shape detection implementation. 
% select whether to proceed with a pure machine learning approach or implement traditional image analysis techniques in their unmanned aerial vehicle (UAV) computer vision system.
\section{Problem Description} 
Performance of this implementation shall be profiled in comparison to a machine learning algorithm implementation to provide information to the AZA club at the University of Arizona.
Analytical data on the performance of the parallelized algorithm will allow the AZA computer vision sub-team to be properly informed when selecting how to proceed with their shape detection implementation. 
The images being analyzed are provided by an unmanned aerial vehicle (UAV) that points a camera to the ground.

% Include and image of one of the pictures that the drone sees here. 

The problem statement requires that we create a GPU-accelerated algorithm to map the image contents to a corresponding shape.
Ideally, a target rate of at least 15 FPS should be reached in order for the program to effectively complete its task at the Student Unmanned Aerial Systems Competition (SUAS) competition.
As preliminary testing has been done, a serial, CPU-based implementation is not capable of achieving such performance, taking a total of greater than 250 seconds to analyze one image with a relatively modern CPU.
Given this knowledge, multiple implementation optimizations must be made in order to achieve sufficient performance. 
While there are many approaches that can be made to optimize the performance of an image detection algorithm, one of the key optimizations that is highly likely to give faster results is implementation of a highly parallelized, GPU-based application of the Hough Transform.
Currently, the implementation done only detects circles in an image, however, AZA wishes to expand this algorithm to include the detection of multiple shapes. 
To detect multiple shapes, the algorithm must increase in complexity. 
Therefore, the algorithm I have currently implemented only detects circles. 
By examining the speedup retrieved by parallelizing the detection of circles, AZA can make an informed decision to elect to continue pursuing further optimizations to detect other shapes as well. 
What follows is an explanation of the Hough Transform algorithm, the way that the algorithm was changed to allow for the detection of circles, and afterwards, an explanation of the implementations done in both the serial domain, and in the parallel domain.
% d the optimizations that were made to improve the detection of circles in an image.

\section{Explanation of the Hough Transform Algorithm}
The Hough Transform argorithm is an image analysis algorithm that allows mapping of shapes on to an image. 
This is done via parameterization of a shape into a formula. 
After a given shape of interest is parameterized, the program then searches the image edge-map and populates a parameter space to identify coordinates of the given shape present in the image. 
When the population of the parameterization is complete, the program can then use the filled paraeter space to populate the image with the shapes that were identified.



\subsection{Image Preprocessing}
Before any sort algorithm that performs shape detection can be conducted, image preprocessing must be completed which conditions the image to allow for proper analysis. 
Assuming that an input is provided as a three color channel image, for example, and RGB image, the image must be converted to grayscale. 
After the image is converted to grayscale, an edge map must be generated. This can be done in multiple ways. 
After an edge map is generated, the program can then run an algorithm to poplate the parameter table, otherwise referred to as the R-Table or the Hough-Array. 




\subsection{$\rho$-$\theta$ Patameterization}
The first implementation of what is known as the modern day Hough Transform parameterized a line onto an image by using a $\rho$-$\theta$ parameterizetion of a line. 
That is, $\rho$ signifies the distance from the specified origin of the image, (which is chosen by the implementor of the algorithm), and $\theta$ specifies the angle of the line from the base angle. 

\begin{equation}
  \rho = x cos(\theta) + y sin(\theta) 
  \label{eq:rho-theta-parameterization}
\end{equation}

Given the equation above, it is possible to map any straight line present in an image using two the two parameters $\rho$ and $\theta$. 

\subsection{Expansion of the Hough Line Transform to Detect Circles}
Modifying the hough transform algorithm to detect circles is a fairly straightforward change. 
Given the equation of a circle: 

\begin{equation}
  r^2 = (x - x_n)^2 + (y - y_n)^2
  \label{eq:circle-parameterization}
\end{equation}
Where $r$, $x_n$, and $y_n$ are the radius, and coordinates of the anchor point of the circle. (In my case, the center of the circle). 
Instead of populating the R-Table with $\rho$ and $\theta$, one can populate the R-Table with three parameters, $r_{n}$, $x_{n}$, and $y_{n}$. 


\subsection{Image Post Processing}
After the R-Table is generated, that is the Hough Transform algorithm execution to completion. 
However, for practical purposes, after the R-Table has been generated, it is necessary to map the shapes that are detected and place them on the image. 
This is usually done via the use of multiple libraries that allow for high level image manipulation. 


\section{Personal Implementation}
Now that the algorithm has been explained, I shal discuss the approach taken to implement the hough transform on my machine. 
The program was coded in C/C++ and the parallelization was implemented in CUDA to take advantage of the GPU on my personal system. 
The only libraries that were used were the standard C/C++ libraries and the Opencv/Opencv-cuda. 
While opencv contains the functions that are required to implement my application, the opencv library was only used to interface with the .PNG image protocol. And placing the detected shapes on the image.
That is, image closing and opening, and drawing shapes on the image.
CMake was used as the build system to simplify compilation accross the systems that are being used in AZA. 

\subsection{Serial Implementation}
Before attempting to make the optimized parallelized version, it was necessary to implement a serilized version that runs on a normal processor. 
For the serial version, there is a 4-level nested for-loop. The algorithm iterates each pixel in both the x-axis and y-axis. Each possible radius is then looped through. The last loop that is included in the implementation is a loop to iterate through every possible theta value, this is done to check the amount of pixels that match the currently parameterized shape.   




























% When using edge-based shape matching algorithms, small shifts in shape orientation of position can greatly in large amounts of edge alignment of shapes\cite{5374408}.
% There are algorithms that can be used to mitigate orientation and position shift errors such as the Hausdorff Distance Algorithms \cite{232073}.
% Regardless, edge-based shape detection alorithms are subject to noise in an image resulting in false edges that make shape detection less accurate because of the sensitivity to noise and illumination changes in a frame \cite{5374408}. 
% Filtering techniques can be implemented to reduce the noise in an image, however implementing those techniques can introduce aliasing into that system \cite{5374408}.




% \section{Ease of Use}

% \subsection{Maintaining the Integrity of the Specifications}

% The IEEEtran class file is used to format your paper and style the text. All margins, 
% column widths, line spaces, and text fonts are prescribed; please do not 
% alter them. You may note peculiarities. For example, the head margin
% measures proportionately more than is customary. This measurement 
% and others are deliberate, using specifications that anticipate your paper 
% as one part of the entire proceedings, and not as an independent document. 
% Please do not revise any of the current designations.

% \section{Prepare Your Paper Before Styling}
% Before you begin to format your paper, first write and save the content as a 
% separate text file. Complete all content and organizational editing before 
% formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
% proofreading, spelling and grammar.

% Keep your text and graphic files separate until after the text has been 
% formatted and styled. Do not number text heads---{\LaTeX} will do that 
% for you.

% \subsection{Abbreviations and Acronyms}\label{AA}
% Define abbreviations and acronyms the first time they are used in the text, 
% even after they have been defined in the abstract. Abbreviations such as 
% IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
% abbreviations in the title or heads unless they are unavoidable.

% \subsection{Units}
% \begin{itemize}
% \item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
% \item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
% \item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
% \item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
% \end{itemize}

% \subsection{Equations}
% Number equations consecutively. To make your 
% equations more compact, you may use the solidus (~/~), the exp function, or 
% appropriate exponents. Italicize Roman symbols for quantities and variables, 
% but not Greek symbols. Use a long dash rather than a hyphen for a minus 
% sign. Punctuate equations with commas or periods when they are part of a 
% sentence, as in:
% \begin{equation}
% a+b=\gamma\label{eq}
% \end{equation}

% Be sure that the 
% symbols in your equation have been defined before or immediately following 
% the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
% the beginning of a sentence: ``Equation \eqref{eq} is . . .''

% \subsection{\LaTeX-Specific Advice}

% Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
% of ``hard'' references (e.g., \verb|(1)|). That will make it possible
% to combine sections, add equations, or change the order of figures or
% citations without having to go through the file line by line.

% Please don't use the \verb|{eqnarray}| equation environment. Use
% \verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
% environment leaves unsightly spaces around relation symbols.

% Please note that the \verb|{subequations}| environment in {\LaTeX}
% will increment the main equation counter even when there are no
% equation numbers displayed. If you forget that, you might write an
% article in which the equation numbers skip from (17) to (20), causing
% the copy editors to wonder if you've discovered a new method of
% counting.

% {\BibTeX} does not work by magic. It doesn't get the bibliographic
% data from thin air but from .bib files. If you use {\BibTeX} to produce a
% bibliography you must send the .bib files. 

% {\LaTeX} can't read your mind. If you assign the same label to a
% subsubsection and a table, you might find that Table I has been cross
% referenced as Table IV-B3. 

% {\LaTeX} does not have precognitive abilities. If you put a
% \verb|\label| command before the command that updates the counter it's
% supposed to be using, the label will pick up the last counter to be
% cross referenced instead. In particular, a \verb|\label| command
% should not go before the caption of a figure or a table.

% Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
% will not stop equation numbers inside \verb|{array}| (there won't be
% any anyway) and it might stop a wanted equation number in the
% surrounding equation.

% \subsection{Some Common Mistakes}\label{SCM}
% \begin{itemize}
% \item The word ``data'' is plural, not singular.
% \item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
% \item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
% \item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
% \item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
% \item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
% \item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
% \item Do not confuse ``imply'' and ``infer''.
% \item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
% \item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
% \item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
% \end{itemize}
% An excellent style manual for science writers is \cite{b7}.

% \subsection{Authors and Affiliations}
% \textbf{The class file is designed for, but not limited to, six authors.} A 
% minimum of one author is required for all conference articles. Author names 
% should be listed starting from left to right and then moving down to the 
% next line. This is the author sequence that will be used in future citations 
% and by indexing services. Names should not be listed in columns nor group by 
% affiliation. Please keep your affiliations as succinct as possible (for 
% example, do not differentiate among departments of the same organization).

% \subsection{Identify the Headings}
% Headings, or heads, are organizational devices that guide the reader through 
% your paper. There are two types: component heads and text heads.

% Component heads identify the different components of your paper and are not 
% topically subordinate to each other. Examples include Acknowledgments and 
% References and, for these, the correct style to use is ``Heading 5''. Use 
% ``figure caption'' for your Figure captions, and ``table head'' for your 
% table title. Run-in heads, such as ``Abstract'', will require you to apply a 
% style (in this case, italic) in addition to the style provided by the drop 
% down menu to differentiate the head from the text.

% Text heads organize the topics on a relational, hierarchical basis. For 
% example, the paper title is the primary text head because all subsequent 
% material relates and elaborates on this one topic. If there are two or more 
% sub-topics, the next level head (uppercase Roman numerals) should be used 
% and, conversely, if there are not at least two sub-topics, then no subheads 
% should be introduced.

% \subsection{Figures and Tables}
% \paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
% bottom of columns. Avoid placing them in the middle of columns. Large 
% figures and tables may span across both columns. Figure captions should be 
% below the figures; table heads should appear above the tables. Insert 
% figures and tables after they are cited in the text. Use the abbreviation 
% ``Fig.~\ref{fig}'', even at the beginning of a sentence.

% \begin{table}[htbp]
% \caption{Table Type Styles}
% \begin{center}
% \begin{tabular}{|c|c|c|c|}
% \hline
% \textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
% \cline{2-4} 
% \textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
% \hline
% copy& More table copy$^{\mathrm{a}}$& &  \\
% \hline
% \multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
% \end{tabular}
% \label{tab1}
% \end{center}
% \end{table}

% \begin{figure}[htbp]
% \caption{Example of a figure caption.}
% \label{fig}
% \end{figure}

% Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
% rather than symbols or abbreviations when writing Figure axis labels to 
% avoid confusing the reader. As an example, write the quantity 
% ``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
% units in the label, present them within parentheses. Do not label axes only 
% with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
% \{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
% quantities and units. For example, write ``Temperature (K)'', not 
% ``Temperature/K''.

% \section*{Acknowledgment}

% The preferred spelling of the word ``acknowledgment'' in America is without 
% an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
% G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
% acknowledgments in the unnumbered footnote on the first page.

% \section*{References}

% Please number citations consecutively within brackets \cite{b1}. The 
% sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
% number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
% the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

% Number footnotes separately in superscripts. Place the actual footnote at 
% the bottom of the column in which it was cited. Do not put footnotes in the 
% abstract or reference list. Use letters for table footnotes.

% Unless there are six authors or more give all authors' names; do not use 
% ``et al.''. Papers that have not been published, even if they have been 
% submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
% that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
% Capitalize only the first word in a paper title, except for proper nouns and 
% element symbols.

% For papers published in translation journals, please give the English 
% citation first, followed by the original foreign-language citation \cite{b6}.

% \begin{thebibliography}{00}
% \bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
% \bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
% \end{thebibliography}
% \vspace{12pt}
% \color{red}
% IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.


% Including the references and creating a bibliography.
\bibliographystyle{ieeetr}
% \bibliographystyle{plain}
\bibliography{refs}


\end{document}

